\section{Approach}

% This should be the most detailed section in general. How did you approach to solve your problem? What worked and what did not work. Here, you should also add your evaluations.

\subsection{Quantum Tape}
\label{sec:approach-quantum-tape}

The main challenge of implementing a brainfuq interpreter lies in the representation of the quantum tape.
In classical brainfuck the tape can be implemented as a plain hash map from indices to tape values.
The indices are signed integers, i.e. they can take on positive and negative values as the pointer is moved left and right.
The tape values are typically bytes representing an 8 bit unsigned integer.

In our implementation we attempt to apply this approach to the quantum tape, which requires some creativity to accomodate the properties of quantum states.
While a naive map from indices to single qubits seems straightforward, it is not sufficient as it can only be used to store product states.
As the gates of the brainfuq specification make up a universal basis of quantum gates, arbitrary and in particular also entangled quantum states can be computed.
To represent arbitrary quantum states we transition to a map $\qtape$ from basis states to complex amplitudes.
Although such a map is capable of storing arbitrary states similar to a plain statevector, the representation of the basis states remains a challenge.
Given that the quantum tape is infinite, i.e. the indices run from $-\infty$ to $\infty$, there are no boundaries that define a reference frame for specifying basis states.
Instead we introduce a $\qmap$ that maps from the indices on the quantum tape to progressively positive qubit indices.
The idea is that when we move the quantum pointer to a new location it is always added on the same side in our classical represntation of the quatum state.
The initial state of our system is $\qtape = \{0: 1.0\}$ and $\qmap = \{0: 0\}$.
When we move the pointer to the left the  $\qmap$ is extended to $\{0: 0, -1: 1\}$, or analogously to $\{0: 0, 1: 1\}$ if we move it right to index 1 on the tape.
Note, that the $\qtape$ remains $\{0: 1.0\}$ for now.
This is because we add new qubits as the most significant bit position, i.e. as $q_n$ in $\ket{q_n ... q_0}$.
For optimization purposes we generally omit the leading zeros in the bitstrings representing the basis states, such that the $\qtape$ does not change when we move the pointer to an unseen index.
As a further optimization we only store those amplitudes in the $\qtape$ map that are not zero.
Apart from incorporating the qubit mapping, applying gates and measuring qubits works equivalently as in standard statevector simulators.

\subsection{Translation from brainfuq to qiskit}
\label{sec:approach-translations}

% translation to openqasm/qiskit

\subsection{Examples}
\label{sec:approach-examples}

To showecasse the functionality of our interpreter and translators, we implemented small instances of some quantum algorithms in the \texttt{examples} folder of the repository.
The script \texttt{examples/ghz.py} (s. Listing \ref{lst:ghz_example}) initializes a $N$-qubit GHZ state.
After applying the Hadamard gate on the first qubit, it leverages classical instructions to iteratively add CNOT gates and moving the quantum pointer.
The number of qubits of the resulting state is controlled by the parameter variable $N$, which is used to set up the loop counter variable that is decremented at the end of each iteration.
For instance $N=5$ produces the brainfuq program \texttt{"\textasciitilde +++++[\#\}*-]"}.
As expected, processing the string puts the quantum tape into the state $\frac{1}{\sqrt{2}} (\ket{00000} + \ket{11111})$, which we confirm by printing the string representation of our quantum simulator.

Further, \texttt{examples/quantum-teleportation.py} implements the circuit for teleporting a qubit through a shared pair of entangled qubits.
In addition to the standard quantum logic it also uses the \texttt{?} instruction to apply Z and X gates based on the classical measurement results.
We initialize q0 as $\ket{-}$ through \texttt{"\textasciitilde ;;;"}, i.e. a Hadamard and 4 consecutive T gates, simulating a Z gate.
We confirm that after the teleportation logic the state of the system features q0 and q1 in their collapsed measurement state and q2 in the $\ket{-}$ state.
Printing the entire system state, we get the \texttt{-0.707 |011⟩ + 0.707 |010⟩} or versions with arbitrary other combinations of the first two qubits.

Lastly, \texttt{examples/grover.py} shows how different software components work together.
To start it translates a qiskit Grover operator, obtained by concatenating a phase oracle and a diffusion operator, into a brainfuq program string.
This string is then embedded into the full brainfuq program as the body to a loop which is executed for the optimal number of Grover iterations which we calculate classically for the given oracle.
We prefix the program with Hadamard gates setting the search qubits into an equal superposition and an optional sufix for measuring the search qubits.
In the script we use an oracle which marks exactly the basis state consisting only of 1-bits, i.e. only one solution.
To prevent this oracle from blowing up in terms of the number of used basis gates we implement a version with $N-2$ auxiliary qubits.
Although this oracle is relatively compact, it still produces brainfuq programs with a significant count of overall instructions.
For $N=6$ for instance the brainfuq program of a single Grover operator is comprised of 21255 classical instruction (including pointer shifts), which results in an overall runtime of almost $30s$. 
Running the script with \texttt{measure=False} and $N=3$ produces the following state where the last qubit is an auxiliary: \texttt{0.972 |1110⟩ - 0.088 |1100⟩ - 0.088 |1000⟩ - 0.088 |0000⟩ - 0.088 |0100⟩ - 0.088 |1010⟩ - 0.088 |0010⟩ - 0.088 |0110⟩}.

\subsection{Edge Case Inputs}
\label{sec:approach-edge-cases}

Given the brainfuq instructions presented in \Cref{sec:background} there are some edge case inputs for which the expected behavior of the system is unclear.
For repeated \texttt{"\#"} without a resetting gate instruction, i.e. \texttt{"*"}, \texttt{"\textasciitilde"} or \texttt{";"}, in between the last \texttt{"\#"} governs which qubit controls the next gate.
If the program does not feature any measurement instructions \texttt{":"} before a \texttt{"?"} the next gate is not executed as the last measurement did not exist and therefore was not $1$ in particular.